---
title: "Vancomycin"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vancomycin}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
This examples shows how to implement the vancomycin population PK model published by Thomson et al (JAC 200, http://www.ncbi.nlm.nih.gov/pubmed/19299472), and sample from the posterior for the individual patients after TDM.

We will first create the Stan model and sample from the posterior. Afterwards, we will
leverage the PKPDsim model in the `pkvancothomson` library to simulate the posterior
for the dependent variable and create a confidence interval plot.

```{r libs, echo=F}
library(PKPDposterior)
library(PKPDsim)
```

First, we define the model and parameters:

```{r params, warning=F, message=F}
parameters <- list(
  CL = 2.99, 
  Q = 2.28, 
  V2 = 0.732, 
  V1 = 0.675,
  TH_CRCL = 0.0154,
  TDM_INIT = 0
)
iiv <- list(
  CL = 0.27,
  Q = 0.49, 
  V1 = 0.15, 
  V2 = 1.3
)
ruv <- list(add = 1.6, prop = 0.15)

model <- new_stan_model(
  parameters = parameters,
  parameter_definitions = list(
    "CL" = "CL * (1.0 + TH_CRCL * ((CRCL[j] * 16.6667) - 66.0))",
    "Q"  = "Q",
    "V1" = "V1 * WT[j]",
    "V2" = "V2 * WT[j]",
    "KA" = "0" 
  ),
  fixed = c("TH_CRCL", "TDM_INIT"),
  covariate_definitions = list(
    "CRCL" = "real", # CrCl in L/hr
    "WT" = "real"    # WT in kg
  ),
  solver = "pmx_solve_twocpt",
  scale = "(V1 * mean(WT))",
  verbose = T
)

model_file <- write_stan_model(model)
```

and compile the model:

```{r compile, warning=F, message=F}
mod <- load_model(
  model_file, 
  force = T,
  verbose = T
)
```

Define the input data and prepare for use in Stan:

```{r data}
regimen <- new_regimen(
  amt = 1500, 
  n = 4, 
  times = c(0, 12, 24, 36), 
  type = 'infusion',
  t_inf = 2
)

covariates <- list(
  WT = new_covariate(value = 70, unit = "kg"),
  CRCL = new_covariate(value = 5, unit = "l/hr")
)

tdm_data <- data.frame(
  t = c(2.5, 11.5), 
  dv = c(40, 14)
)

data <- new_stan_data(
  regimen,
  covariates, 
  tdm_data,
  dose_cmt = 2,
  parameters = parameters,
  fixed = c("TH_CRCL", "TDM_INIT"),
  iiv = iiv,
  ruv = list(
    prop = 0.15,
    add = 1.6
  ),
  ltbs = FALSE
)
```

Then, we can sample from the posterior using:

```{r sample, message=F, warning=F, error=F}
post <- get_mcmc_posterior(
  mod = mod,
  data = data
)
```

Summary info about the posterior:

```{r post-info, warning=F}
post
```

Plot parameter distributions:

```{r param-dist, warning=F, message=F}
plot_params(post)
```

We can now generate the posterior for the dependent value using the PKPDsim model.

First we define an additional required covariates and a mapping between parameters
in the PKPDposterior model and the PKPDsim model.

```{r mapping}
mapping <- list("V1" = "V") # mapping between parameters in the 
covariates$CL_HEMO <- new_covariate(0) # required covariate for PKPDsim model
```

The PKPDsim model package `pkvancothomson` that is used in this example can be 
installed using:

```{r vanco-pkg, eval=F}
model_from_api(
  system.file("models", "pk_vanco_thomson.json5", package = "PKPDsim"),
  to_package = TRUE,
  install_all = TRUE
)
```

Then we can simulate the posterior of the dependent variable: 

```{r pred-posterior, warning=F, message=F}
library(pkvancothomson)
pred_post <- sim_from_draws(
  post, 
  model = pkvancothomson::model(),
  map = mapping,
  regimen = regimen,
  covariates = covariates,
  n = 200,
  summarize = TRUE
)
pred_prior <- sim_from_draws(
  post, 
  model = pkvancothomson::model(), 
  map = mapping,
  regimen = regimen,
  covariates = covariates,
  prior = TRUE,
  n = 200,
  summarize = TRUE
)
```

Now we can plot the confidence interval for the posterior and prior predictions using:

```{r plot-ci}
plot_predictions(
  pred_prior, 
  obs = tdm_data
)
plot_predictions(
  pred_post, 
  obs = tdm_data
)
```
