---
title: "Neutropenia (PKPD)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Neutropenia (PKPD)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
This examples shows how to implement the Friberg neutropenia model. What is implemented here
is a generic version of it, i.e. not for a specific drug. We will sample from the posterior for the individual patients using TDM data and neutrophil count data.

We will first create the Stan model and sample from the posterior. 

```{r libs, echo=F}
library(PKPDposterior)
library(PKPDsim)
```

First, we define the model and parameters. Note that in this exmaple we specify `ruv` 
as a list that contains separate residual errors for `pk`- and `pd`-labeled observations.
  
```{r params, warning=F, message=F}
parameters <- list(CL = 5, V = 50,
                   SLOPE = 0.1, MTT = 100, 
                   CIRC0 = 5, GAMMA = 0.2)
iiv <- list(
  CL = 0.2, V = 0.5, 
  SLOPE = 1.0, MTT = 0.2, 
  CIRC0 = 0.5, GAMMA = 0.2
)
ruv <- list(
  pk = list(add = 0.2),
  pd = list(add = 0.3)
)
parameter_definitions <- list(
  "CL" = "CL",
  "Q" = 0,
  "V" = "V",
  "V2" = 1,
  "KA" = 0,
  "MTT" = "MTT",
  "CIRC0" = "CIRC0",
  "GAMMA" = "GAMMA",
  "ALPHA" = "SLOPE"
)
ode <- "
  real ka  = 0;
  real k10 = CL / V;
  real k12 = 0;
  real k21 = 0;
  real ktr = 4 / MTT;
  
  real conc;
  real EDrug;
  real transit1;
  real transit2;
  real transit3;
  real circ;
  real prol;
  
  dAdt[1] = -KA * A[1];
  dAdt[2] =  KA * A[1] - (k10 + k12) * A[2] + k21 * A[3];
  dAdt[3] = k12 * A[2] - k21 * A[3];
  conc = A[2] / V;
  
  EDrug = ALPHA * conc; // slope model, not Emax
  prol = A[4] + CIRC0;
  transit1 = A[5] + CIRC0;
  transit2 = A[6] + CIRC0;
  transit3 = A[7] + CIRC0;
  circ = fmax(machine_precision(), A[8] + CIRC0); // Device for implementing a modeled 
  
  // initial condition
  dAdt[4] = ktr * prol * ((1 - EDrug) * ((CIRC0 / circ)^GAMMA) - 1);
  dAdt[5] = ktr * (prol - transit1);
  dAdt[6] = ktr * (transit1 - transit2);
  dAdt[7] = ktr * (transit2 - transit3);
  dAdt[8] = ktr * (transit3 - circ);
"

model <- new_stan_model(
  parameters = parameters,
  parameter_definitions = parameter_definitions,
  ode = ode,
  covariate_definitions = NULL,
  solver = 'pmx_solve_rk45',
  obs_types = c("pk", "pd"),
  custom_ipred = list(
    "pk" = "A[2, ] ./ V;",
    "pd" = "A[8, ] + theta[7];"
  ),
  verbose = T
)
model_file <- write_stan_model(model)
```

and compile the model:
  
```{r compile, warning=F, message=F}
mod <- load_model(
  model_file
)
```

Define the input data and prepare for use in Stan. Note that in the definition of 
`tdm_data` we use the `type` argument to distinguish between PK and PD observations.
  
```{r data}
regimen <- new_regimen(
  amt = 1500, 
  n = 3, 
  times = c(0, 24, 48), 
  type = 'infusion',
  cmt = 2,
  t_inf = 2
)
covariates <- NULL
tdm_data <- data.frame(
  t = c(2.5, 11.5), 
  dv = c(40, 14),
  type = "pk"
)
pd_data <- data.frame(
  t = c(3, 6, 9, 12) * 24, 
  dv = c(5, 1.5, .8, 2),
  type = "pd"
)
comb_data <- rbind(
  tdm_data, 
  pd_data
)

data <- new_stan_data(
  regimen,
  covariates = covariates, 
  data = comb_data,
  parameters = parameters,
  iiv = iiv,
  ltbs = list(pk = TRUE, pd = TRUE),
  ruv = ruv,
  dose_cmt = 2
)
```

Then, we can sample from the posterior using:
  
```{r sample, message=F, warning=F, error=F}
post <- get_mcmc_posterior(
  mod = mod,
  data = data
)
```

Summary info about the posterior:
  
```{r post-info, warning=F}
post
```

Plot parameter distributions:
  
```{r param-dist, warning=F, message=F}
plot_params(post)
```
